An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2016-10-12 aistats2017.tex
More shortening of paper, roughly 8.25 pages

We now derive error bounds for our M-H test, and for the approximate target
distribution that it generates. From Table~\ref{tab:xcorr}, we know that it is
possible to generate the correction samples $X_{\rm corr}$ with a CDF error
approaching single-precision floating point error. We therefore treat $X_{\rm
corr}$ as a sample from the exact correction distribution and we will not
analyze its errors.

In the most similar prior works,~\citet{cutting_mh_2014} uses asymptotic
arguments based on the CLT to argue that its approximate acceptance test error
tends to zero as batch size increases, but no quantitative bounds are given.
In~\citet{icml2014c1_bardenet14}, explicit bounds are given, but they depend on
bounding:
\begin{equation}\label{eq:bad_bound}
    C_{\theta, \theta'} = \max_{1\leq i\leq N}|\log p(x_i\mid\theta') - \log p(x_i\mid\theta)|.
\end{equation}
Such bounds can be derived efficiently for models such as logistic regression,
but it is unclear how to derive them for a complex model such as a neural
network. In general, since a new $\theta'$ value is obtained each iteration, one
would need to use all the $p(x_i\mid \theta')$ terms\footnote{The sample code
provided by the authors of~\citet{icml2014c1_bardenet14} in fact computes
$C_{\theta,\theta'}$ explicitly for each sample generated, i.e. it traverses the
entire data, thus providing no performance advantage over the complete test.}.
In contrast, we use quantitative forms of the Central Limit Theorem which rely
on measurable statistics from a single minibatch. Thus a sampler using our
approach does not need to see data beyond the current minibatch.

@@

\begin{corollary}\label{cor:our_bound_delta_prime}
We can now substitute into Equation~(\ref{eq:clt-bounds2}) and displace by the mean, giving:
\begin{align*}
    \sup_y\left|{\rm Pr}(\Delta^*< y) - \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)\right| &\leq \frac{6.4E|X|^3+2E|X|}{\sqrt{b}} \\
    &= \epsilon(\theta,\theta',b).
\end{align*}
\end{corollary} %\vspace{-10pt}
Corollary~\ref{cor:our_bound_delta_prime} shows that the distribution of
$\Delta^*$ approximates a Normal distribution with mean $\Delta$ and variance
$s^2_{\Delta^*}$. Furthermore, it bounds the error with \emph{estimable
quantities}: both $E|X|$ and $E|X|^3$ can be estimated as means of $|\Lambda_i
- \Lambda|$ and $|\Lambda_i - \Lambda|^3$, respectively, on each minibatch. We
expect this will often be accurate enough on minibatches with hundreds or
thousands of points, but otherwise bootstrap CIs can be computed from those
sequences. Since the bounds are monotone in $E|X|$ and $E|X|^3$, using upper
bootstrap CI limits will provide high-confidence error bounds.

>>>>

We now derive error bounds for our M-H test and the approximate target
distribution it generates. (We treat $X_{\rm corr}$ samples from the exact
correction distribution based on the results in Table~\ref{tab:xcorr}.)

In the most similar prior works,~\citet{cutting_mh_2014} uses asymptotic
arguments based on the CLT to argue that its approximate acceptance test error
tends to zero as batch size increases, but no quantitative bounds are given.
In~\citet{icml2014c1_bardenet14}, explicit bounds are given, but they depend on
bounding:
\begin{equation}\label{eq:bad_bound}
    C_{\theta, \theta'} = \max_{1\leq i\leq N}|\log p(x_i\mid\theta') - \log p(x_i\mid\theta)|.
\end{equation}
It is possible to analytically determine $C_{\theta,\theta'}$ for logistic
regression, but it is unclear how to do so for more complicated models, and one
generally needs to use all $p(x_i\mid \theta')$ terms\footnote{The sample code
provided by the authors of~\citet{icml2014c1_bardenet14} in fact computes
$C_{\theta,\theta'}$ explicitly for each sample generated, i.e. it traverses the
entire data, thus providing no performance advantage over the complete test.} to
get $C_{\theta,\theta'}$.  In contrast, we use quantitative forms of the Central
Limit Theorem which rely on measurable statistics from a single minibatch. Thus
a sampler using our approach does not need to see data beyond the current
minibatch.

@@

\begin{corollary}\label{cor:our_bound_delta_prime}
We can now substitute into Equation~(\ref{eq:clt-bounds2}) and displace by the mean, giving:
\begin{align*}
    \sup_y\left|{\rm Pr}(\Delta^*< y) - \Phi\left(\frac{y-\Delta}{s_{\Delta^*}}\right)\right| &\leq \frac{6.4E|X|^3+2E|X|}{\sqrt{b}} \\
    &= \epsilon(\theta,\theta',b).
\end{align*}
\end{corollary} %\vspace{-10pt}
Corollary~\ref{cor:our_bound_delta_prime} shows that the distribution of
$\Delta^*$ approximates a Normal distribution with mean $\Delta$ and variance
$s^2_{\Delta^*}$. Furthermore, it bounds the error with \emph{estimable
quantities}: both $E|X|$ and $E|X|^3$ can be estimated as means of $|\Lambda_i
- \Lambda|$ and $|\Lambda_i - \Lambda|^3$, respectively, on each minibatch. We
expect this will often be accurate enough on minibatches with hundreds or
thousands of points, but otherwise bootstrap CIs can be computed from those
sequences.

% Daniel: I removed this sentence to save space.
% Since the bounds are monotone in $E|X|$ and $E|X|^3$, using upper
% bootstrap CI limits will provide high-confidence error bounds.
